---
title: "MultiView.LOCUS Analysis for Three Simulation Datasets"
output:
  html_document:
    toc: true
    toc_float: true
---

### The Rmd File is self-contained, no need to download data files separately. 

## Install the package

```{r setup, include=T}
if (!requireNamespace("devtools", quietly = TRUE))
  install.packages("devtools")
if (!requireNamespace("MultiView.LOCUS", quietly = TRUE)) {
  devtools::install_github("Emory-CBIS/MultiView.LOCUS")
}# Additional packages used in this report
for (pkg in c("lattice", "gridExtra", "grid")) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}

library(MultiView.LOCUS)
library(lattice)
library(gridExtra)
library(grid)
```

## (1) Helper functions - hidden

**Functions for plotting and parameter selections**

```{r helper-functions, include=FALSE}
plot_mat <- function(mat, limits = 5) {
  mat[mat > limits] <- limits
  mat[mat < -limits] <- -limits
  levelplot(
    mat,
    col.regions = colorRampPalette(c("blue", "white", "red")),
    at = do.breaks(c(-limits, limits), 100),
    scales = list(x = list(at = NULL), y = list(at = NULL), tck = c(0, 0)),
    xlab = "", ylab = ""
  )
}

plot_conn_wrapper <- function(vec, V) {
  mat <- Ltrinv(vec, V, d = FALSE)
  plot_conn(mat)
}

plot_mat_component <- function(res, view_idx, k, V, scale = 6, limits = 5) {
  vec <- res$S[[view_idx]][k, ]
  if (max(abs(vec)) == 0) {
    return(plot_mat(Ltrinv(vec, V, FALSE), limits = limits))
  }
  plot_mat(scale * Ltrinv(vec / max(abs(vec)), V, FALSE), limits = limits)
}

visualize_plot_mat_grid <- function(res, V, q_common, q_specific, scale = 6, limits = 5) {
  n_total <- q_common + q_specific

  view1 <- lapply(seq_len(n_total), function(k) {
    grid::grid.grabExpr(print(plot_mat_component(res, 1, k, V, scale, limits), newpage = FALSE))
  })
  view2 <- lapply(seq_len(n_total), function(k) {
    grid::grid.grabExpr(print(plot_mat_component(res, 2, k, V, scale, limits), newpage = FALSE))
  })

  plot_list <- c(view1, view2)
  col_titles <- c(paste0("Common ", seq_len(q_common)),
                  paste0("Specific ", seq_len(q_specific)))
  col_grobs <- lapply(col_titles, function(title) {
    textGrob(title, gp = gpar(fontsize = 11, fontface = "bold"))
  })

  row_grobs <- lapply(c("View 1", "View 2"), function(title) {
    textGrob(title, rot = 90, gp = gpar(fontsize = 12, fontface = "bold"))
  })

  plot_grid <- arrangeGrob(grobs = plot_list, nrow = 2, ncol = n_total)
  arrangeGrob(
    plot_grid,
    left = arrangeGrob(grobs = row_grobs, ncol = 1),
    bottom = arrangeGrob(grobs = col_grobs, nrow = 1),
    padding = unit(2, "lines")
  )
}

visualize_common_specific <- function(res, V, q_common, q_specific) {
  n_total <- q_common + q_specific
  plot_list <- vector("list", length = n_total * 2)

  for (i in seq_len(n_total)) {
    plot_list[[i]] <- plot_conn_wrapper((res$S[[1]][i, ])*10, V)
  }
  for (i in seq_len(n_total)) {
    plot_list[[n_total + i]] <- plot_conn_wrapper((res$S[[2]][i, ])*10, V)
  }

  col_titles <- c(paste0("Common ", seq_len(q_common)),
                  paste0("Specific ", seq_len(q_specific)))
  col_grobs <- lapply(col_titles, function(title) {
    textGrob(title, gp = gpar(fontsize = 11, fontface = "bold"))
  })

  row_grobs <- lapply(c("View 1", "View 2"), function(title) {
    textGrob(title, rot = 90, gp = gpar(fontsize = 12, fontface = "bold"))
  })

  plot_grid <- arrangeGrob(grobs = plot_list, nrow = 2, ncol = n_total)
  arrangeGrob(
    plot_grid,
    left = arrangeGrob(grobs = row_grobs, ncol = 1),
    bottom = arrangeGrob(grobs = col_grobs, nrow = 1),
    padding = unit(2, "lines")
  )
}

compute_bic_grid <- function(Data_x, Data_y, q_common, q, V,
                             gamma, rho, espli1, espli2,
                             phi_seq = seq(1, 3, by = 0.5),
                             psi_seq = seq(0.1, 1, by = 0.3),
                             penalty = "SCAD") {
  BIC_matrix <- matrix(
    NA,
    nrow = length(phi_seq),
    ncol = length(psi_seq),
    dimnames = list(paste0("phi=", phi_seq), paste0("psi=", psi_seq))
  )

  for (i in seq_along(phi_seq)) {
    for (j in seq_along(psi_seq)) {
      cat("Running phi =", phi_seq[i], "and psi =", psi_seq[j], "\n")

      score <- tryCatch({
        result_bic <- multi_view_decomposition(
          Y = list(Data_x, Data_y),
          q_common = q_common,
          q = q,
          V = V,
          penalty = penalty,
          phi = phi_seq[i],
          psi = psi_seq[j],
          gamma = gamma,
          rho = rho,
          espli1 = espli1,
          espli2 = espli2,
          MaxIteration = 500,
          silent = TRUE
        )

        bic_values <- calculate_bic(Y = list(Data_x, Data_y), res = result_bic)
        if (!all(is.finite(bic_values))) {
          NA_real_
        } else {
          sum(bic_values)
        }
      }, error = function(e) {
        NA_real_
      })

      BIC_matrix[i, j] <- score
    }
  }

  if (all(is.na(BIC_matrix))) {
    optimal_phi <- NA_real_
    optimal_psi <- NA_real_
  } else {
    optimal_indices <- which(BIC_matrix == min(BIC_matrix, na.rm = TRUE), arr.ind = TRUE)
    optimal_phi <- phi_seq[optimal_indices[1]]
    optimal_psi <- psi_seq[optimal_indices[2]]
  }

  list(
    BIC_matrix = BIC_matrix,
    phi_seq = phi_seq,
    psi_seq = psi_seq,
    optimal_phi = optimal_phi,
    optimal_psi = optimal_psi
  )
}
```

## (2) Data Loading

This report runs the same tuning + final model workflow for:

- **simu_1**: `Data_x_simu1.rds` + `Data_y_simu1.rds`
- **simu_2**: `Data_x_simu2.rds` + `Data_y_simu2.rds`
- **simu_3**: `Data_x.rds` + `Data_y.rds`

```{r load-data}
load_rds_from_sources <- function(filename,
                                  package = "MultiView.LOCUS",
                                  urls = character(0),
                                  local_fallback = character(0)) {
  # 1) Installed package extdata
  pkg_file <- system.file("extdata", filename, package = package)
  if (nzchar(pkg_file) && file.exists(pkg_file)) {
    return(readRDS(pkg_file))
  }

  # 2) URL(s)
  for (u in urls) {
    obj <- tryCatch({
      con <- url(u, open = "rb")
      on.exit(close(con), add = TRUE)
      readRDS(con)
    }, error = function(e) NULL)

    if (!is.null(obj)) {
      return(obj)
    }
  }

  # 3) Local fallback path(s)
  for (p in local_fallback) {
    if (file.exists(p)) {
      return(readRDS(p))
    }
  }

  stop("Could not load ", filename,
       " from package extdata, URL, or local fallback paths.")
}

repo_raw <- "https://raw.githubusercontent.com/Emory-CBIS/MultiView.LOCUS/main"

sim_data <- list(
  simu_1 = list(
    Data_x = load_rds_from_sources(
      "Data_x_simu1.rds",
      urls = c(file.path(repo_raw, "inst/extdata/Data_x_simu1.rds")),
      local_fallback = c("example_data/Data_x_simu1.rds")
    ),
    Data_y = load_rds_from_sources(
      "Data_y_simu1.rds",
      urls = c(file.path(repo_raw, "inst/extdata/Data_y_simu1.rds")),
      local_fallback = c("example_data/Data_y_simu1.rds")
    )
  ),
  simu_2 = list(
    Data_x = load_rds_from_sources(
      "Data_x_simu2.rds",
      urls = c(file.path(repo_raw, "inst/extdata/Data_x_simu2.rds")),
      local_fallback = c("example_data/Data_x_simu2.rds")
    ),
    Data_y = load_rds_from_sources(
      "Data_y_simu2.rds",
      urls = c(file.path(repo_raw, "inst/extdata/Data_y_simu2.rds")),
      local_fallback = c("example_data/Data_y_simu2.rds")
    )
  ),
  simu_3 = list(
    Data_x = load_rds_from_sources(
      "Data_x.rds",
      urls = c(file.path(repo_raw, "example_data/Data_x.rds")),
      local_fallback = c("example_data/Data_x.rds")
    ),
    Data_y = load_rds_from_sources(
      "Data_y.rds",
      urls = c(file.path(repo_raw, "example_data/Data_y.rds")),
      local_fallback = c("example_data/Data_y.rds")
    )
  )
)

sim_data <- lapply(sim_data, function(d) {
  d$Data <- list(d$Data_x, d$Data_y)
  d
})

lapply(sim_data, function(d) c(dim(d$Data_x), dim(d$Data_y)))
```


# Simulation Study 1 (`simu_1`) — tune phi/psi then final result

```{r simu1-tune,warning=FALSE}
Data_x <- sim_data$simu_1$Data_x
Data_y <- sim_data$simu_1$Data_y
Data <- list(Data_x, Data_y)

simu_1_tuning <- compute_bic_grid(
  Data_x = Data_x,
  Data_y = Data_y,
  q_common = 3,
  q = c(2, 2),
  V = 50,
  gamma = 3,
  rho = 0.95,
  espli1 = 1e-3,
  espli2 = 1e-3,
  phi_seq = seq(1, 3, by = 0.5),
  psi_seq = seq(0.1, 1, by = 0.3)
)

simu_1_tuning$BIC_matrix
simu_1_tuning$optimal_phi
simu_1_tuning$optimal_psi
```

```{r simu1-final}
simu_1_res <- multi_view_decomposition(
  Y = Data,
  q = c(2, 2),
  q_common = 3,
  V = 50,
  penalty = "SCAD",
  phi = simu_1_tuning$optimal_phi,
  psi = simu_1_tuning$optimal_psi,
  gamma = 3,
  rho = 0.95,
  espli1 = 1e-3,
  espli2 = 1e-3,
  MaxIteration = 500,
  silent =T
)

```

```{r simu1-visual, fig.width=16, fig.height=7}
# simu_1 uses plot_mat-based grid: components 1-3 are common, 4-5 are specific
grid.newpage()
grid.draw(visualize_plot_mat_grid(simu_1_res, V = 50, q_common = 3, q_specific = 2,scale = 4))
```

# Simulation Study 2 (`simu_2`) — tune phi/psi then final result

```{r simu2-tune,warning=FALSE}
Data_x <- sim_data$simu_2$Data_x
Data_y <- sim_data$simu_2$Data_y
Data <- list(Data_x, Data_y)

simu_2_tuning <- compute_bic_grid(
  Data_x = Data_x,
  Data_y = Data_y,
  q_common = 2,
  q = c(1, 1),
  V = 50,
  gamma = 3,
  rho = 0.98,
  espli1 = 1e-3,
  espli2 = 1e-3,
  phi_seq = seq(1, 3, by = 0.5),
  psi_seq = seq(0.1, 1, by = 0.3)
)

simu_2_tuning$BIC_matrix
simu_2_tuning$optimal_phi
simu_2_tuning$optimal_psi
```

```{r simu2-final}
simu_2_res <- multi_view_decomposition(
  Y = Data,
  q = c(1, 1),
  q_common = 2,
  V = 50,
  penalty = "SCAD",
  phi = simu_2_tuning$optimal_phi,
  psi = simu_2_tuning$optimal_psi,
  gamma = 3,
  rho = 0.98,
  espli1 = 1e-3,
  espli2 = 1e-3,
  MaxIteration = 5000,
  silent = T
)

```

```{r simu2-visual, fig.width=14, fig.height=7}
# simu_2 uses plot_mat-based grid: components 1-2 are common, 3 is specific
grid.newpage()
grid.draw(visualize_plot_mat_grid(simu_2_res, V = 50, q_common = 2, q_specific = 1,scale = 6))
```

# Simulation Study 3 (`simu_3`) — show results only (no tuning)

```{r simu3-final}
Data_x <- sim_data$simu_3$Data_x
Data_y <- sim_data$simu_3$Data_y
Data <- list(Data_x, Data_y)

simu_3_res <- multi_view_decomposition(
  Y = list(Data_x, Data_y),
  q_common = 2,
  q = c(2, 2),
  V = 264,
  penalty = "L1",
  phi = 0.5,
  psi = 0.5,
  gamma = 3,
  rho = 0.95,
  espli1 = 1e-3,
  espli2 = 1e-3,
  MaxIteration = 500,
  silent = T
)

```

```{r simu3-visual, fig.width=16, fig.height=7}
# simu_3 annotation: components 1-2 are common, 3-4 are specific
grid.newpage()
grid.draw(visualize_common_specific(simu_3_res, V = 264, q_common = 2, q_specific = 2))
```